Apparent preferences for cognitive effort fade when multiple forms of effort

### 2020-11-13 Cost_manuscript.Rmd
reorganization of figures

---
title: Apparent preferences for cognitive effort fade when multiple forms of effort
  and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey^1^2, Gary Kane^1^2, & Joseph T. McGuire^1^2"
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
- \usepackage{indentfirst}
output:
  pdf_document:
    highlight: haddock
    keep_tex: no
  word_document: #default
    reference_docx: Style_reference_draft.docx
  html_document:
    df_print: paged
csl: apa.csl
indent: true
#subtitle: 'Abbreviated Title: Heterogeneity in the functional topography of mPFC'
bibliography: cost_paper.bib
---

@@

```{r Trial plot (Fig. 1), out.width = "80%", fig.align = "center", fig.cap= "General foraging trial structure. On each trial, participants were offered to earn money (4, 8, or 20 cents) by sustaining effort or waiting during the handling time (2, 10, or 14 s). The end of a trial was followed by a travel time (handling and travel times always added up to 16 s). Participants could skip unfavorable trials and immediately start traveling to a potentially better offer. In the between-subject experiment, cost was fixed per participant and handling time varied per block. In the within-subject version, handling time was fixed at 10 s, but a combination of effort and delay trials changed per block. Possible costs, handling times, and rewards were fully disclosed in order to avoid experience-dependent learning."}

@@

  where *R* and *p* are the reward, and acceptance probability of offer i, respectively, *H* is the handling time for the block, and *T* its associated travel time. This gave us the total reward per second attainable in each block as a function of the acceptance threshold (i.e. the lowest amount accepted). Figure 2 shows possible earnings for each choice strategy, as well as the lowest amount participants should accept in order to maximize their rewards (circled dots). For example, a participant in a 10 s handling block should accept 8-cent and 20-cent rewards (and reject 4-cent rewards) to maximize their reward rate. Note that accepting every offer was often detrimental to participant earnings. Significant deviations from the optimal strategy can be understood as reflecting changes in subjective OC. This approach allowed us to test if certain demands boosted the value of offers, expressed as a subjective OC that fell below the optimal rate.

```{r Optimal plot (Fig. 2), out.width = "80%", fig.align = "center", fig.cap= "Possible earnings per second for each acceptance threshold (i.e. the smallest amount accepted) for each handling time. Circles denote the reward-maximizing threshold for each block, which is described in the table. Experiments shared the 10 s handling block, in which it was optimal to skip all 4 cent offers."}

@@

## MAYBE BETTER WAY TO DO PAIRWISE COMPARISONS, USING VECTORIZE FROM BELOW
# OCanova$perms <- summaryOC$all %>%
#   select(Cost, Gamma) %>%
#   plyr::dlply("Cost", identity) %>%
#   do.call(cbind, .) %>%
#   select_if(is.numeric) %>%
#   rename(Physical = Physical.Gamma,
#          Cognitive = Cognitive.Gamma,
#          Wait = Wait.Gamma,
#          Easy = Easy.Gamma) %>%
#   outer(., ., FUN = Vectorize(permute), simple = T)
# diag(OCanova$perms) <- 1

@@

```{r General prop & earnings plots (Fig. 3), fig.width = 10, fig.height = 4, fig.align = "center", fig.cap = "Left: Proportion accepted per cost. Middle: Total number of dollars earned by each group by the end of the experiment. Right: The relationship between proportion accepted and total earned. Consistent with the foraging design, participants who over and under accepted earned the least."}
# plots generated in 16.2.1. and 16.1.3.
 prop1 | earnings1 | earnFitplot_btw
```

@@

```{r mixed effects plots (Fig. 4), out.width = "80%", fig.align = "center", fig.cap = "Comparing acceptance rates among costs for each reward and handling combination. A: Proportion accepted by each group for every combination of handling time and reward. Grey dots indicate the reward-maximizing acceptance for each combination. B: Matrix portraying the coefficients (color scale) and significance that resulted from switching the reference cost condition (crosses mark non-significant comparisons). Each entry shows how much more likely a group was to accept compared to the reference condition (row). C: AIC and BIC values for a forward mixed-effects model comparison, ranging from a base intercept-only to a full-interaction model."}
#hrPlot_btw / (ABIC_plot_btw | ME_matrix_btw) # doesn't work with corrplot.. oh well
include_graphics("./Images/FIG4.png")
```

@@

```{r Survival and prepost plots (Fig. 5), fig.width = 8, fig.height = 7, fig.align = "center", fig.cap = "Top: Survival curve (left), and distribution of times when participants quit a trial over time. Bottom: Proportion accepted in the second half as a function of first half acceptances (left). Dashed lines indicate the mean for each group. On the right, the proportion accepted over time per group. The order of blocks was repeated in the second half (indicated by the black dashed line)."}
# plots generated in 16.3.1. and 16.3.2.
(suppressWarnings(survPlot_btw) | RTplot_btw) / (prepost_btw$scatterplot | prepost_btw$propBlocks)
```

@@

```{r Prepost mixed effects plot (Fig. 6), fig.align = "center", fig.cap = "Distilling acceptance behavior before and after the break. Left: Proportion of acceptances per handling, reward, and half, showing that effortful demands became increasingly aversive over time. Right: Mixed-effects coefficients denoting the comparison among costs acceptances for each half separately (mirrored along the diagonal). Even though acceptances steadily decreased in the effort groups, the relative preferences were preserved."}
include_graphics("./Images/FIG6.png")

@@

  Analyses from the previous experiment were repurposed to examine cost differences within individuals. We started by testing the hypothesis that participants would successfully integrate the statistics of the environment. A mixed-effects logistic regression once again showed that larger reward amounts significantly increased acceptance probabilities ($\beta$ = `r historyCoeffs_wth["Offer", "Estimate"]`, SE = `r historyCoeffs_wth["Offer", "Std. Error"]`, *p* `r report_p(historyCoeffs_wth["Offer", "Pr(>|z|)"])`), and that participants became more selective if they had recently observed large offers ($\beta$ = `r historyCoeffs_wth["recentOffers", "Estimate"]`, SE = `r historyCoeffs_wth["recentOffers", "Std. Error"]`, *p* `r report_p(historyCoeffs_wth["recentOffers", "Pr(>|z|)"])`). Model comparisons suggested that participants were influenced by the `r historyIndex_wth + 1` most recent rewards (AIC = `r round(AIC(historySize_wth[[historyIndex_wth]]), digits = 2)`). However, this time participants were unaffected by recent reward misses (*p* > 0.05).
  As before, a linear model predicting earnings showed that participants who completed too few or too many trials earned the least, in line with the experimental design (Figure 8A; general linear model with quadratic term, F = `r round(summary(lmEarn)$fstatistic[1], digits = 2)`, $\beta$ = `r round(summary(lmEarn)$coefficients[3,1], digits= 2)`, SE = `r round(summary(lmEarn)$coefficients[3, 2], digits = 2)`, $R^2$ = `r round(summary(lmEarn)$r.squared, digits = 2)`, *p* `r report_p(summary(lmEarn)$coefficients[3, 4])`). When assessing the optimality of acceptances for 4 and 8 cents (99% of 20 cent offers were correctly accepted), chi-squared tests against the optimal rate of 50% showed that participants under-accepted physical effort trials (proportion accepted = `r round(summary_costvsreward["Physical", "probability"], digits = 2)`, $\chi^2$ = `r round(summary_costvsreward["Physical", "chisquared"], digits = 2)`, *p* `r report_p(summary_costvsreward["Physical", "pval"])`; all other *p* > 0.05). Despite this deviation, participants earned a similar total amount per cost on average (mean Cognitive = `r filter(earnings_within, Cost == "Cognitive")$mEarn`, SD = `r filter(earnings_within, Cost == "Cognitive")$sdEarn`; mean Wait-C = `r filter(earnings_within, Cost == "Wait-C")$mEarn`, SD = `r filter(earnings_within, Cost == "Wait-C")$sdEarn`; mean Physical = `r filter(earnings_within, Cost == "Physical")$mEarn`, SD = `r filter(earnings_within, Cost == "Physical")$sdEarn`; mean Wait-P = `r filter(earnings_within, Cost == "Wait-P")$mEarn`, SD = `r filter(earnings_within, Cost == "Wait-P")$sdEarn`). These results convey that participants in both experiments were similarly influenced by the characteristics of the environment, although participants in the present experiment gravitated more towards optimality.

@@

```{r Earnings, prop completed, ME plots (Fig. 7), fig.width = 8, fig.height = 7, fig.align = "center", fig.cap = "A: Quadratic relationship between trial completions and total earnings. B: proportion of acceptances of each reward per cost condition. C: Matrix of mixed-effects coefficients showing comparisons among costs (while controlling for reward amount). Comparisons between effort and wait trials from different blocks are omitted, as these are not considered in our original predictions. Crosses mark non-significant comparisons. D: AIC and BIC values for models of increasing complexity (a priori = AllMain)."}

@@

  Next, we addressed the hypothesis that passive waiting trials would be less costly than effortful ones. Even though the reward acceptance pattern at 10 s handling was similar to the between-subjects experiment (Figure 8B), differences among cost conditions were less evident. Moreover, these differences contradicted our predictions, as cognitive effort remained the least costly among demands. A mixed effects logistic regression confirmed once again that increases in offer amount made acceptances significantly more likely (global intercept = `r round(fixef(mixLogis_main$Cognitive)["(Intercept)"], digits = 2)`, $\beta$ = `r round(fixef(mixLogis_main$Cognitive)["Offer"], digits = 2)`, SE = `r round(summary(mixLogis_main$Cognitive)[[10]]["Offer", "Std. Error"], digits = 2)`, *p* `r report_p(summary(mixLogis_main$Cognitive)[[10]]["Offer", "Pr(>|z|)"])`), but that the only significant difference was the higher likelihood to accept cognitive offers compared to physical ones (Figure 7C; $\beta$ = `r round(mixSummary_wth$Betas["Physical", "Cognitive"], digits = 2)`, *p* `r report_p(mixSummary_wth$Pvals["Physical", "Cognitive"])`). Model comparisons suggested that even though a model with all main effects explained a large portion of the variance ($R^2$ = `r round(mixLogis_compare$Rsq, digits = 2)`), there were similar AIC and BIC values for any model including reward amount (Figure 8D). An analysis of deviance among the three models showed that adding cost condition to reward amount significantly improved fit (AIC~reward~ = `r filter(abicLogis_compare, Model == "Reward")$AIC`, AIC~a-priori~ = `r filter(abicLogis_compare, Model == "AllMain")$AIC`, $\chi^2$(`r mixLogis_compare$aod[[7]][2]`) = `r round(mixLogis_compare$aod$Chisq[2], digits = 2)`, *p* `r report_p(mixLogis_compare$aod[[8]][2])`). However, probing the interaction between these features significantly improved fits from the main effects model (AIC~Interaction~ = `r filter(abicLogis_compare, Model == "Interaction")$AIC`, $\chi^2$(`r mixLogis_compare$aod[[7]][3]`) = `r round(mixLogis_compare$aod$Chisq[3], digits = 2)`, *p* `r report_p(mixLogis_compare$aod[[8]][3])`). (**discuss with Joe: the interaction flips results over, such that both waits > pysical effort = cognitive effort...as predicted originally? Can't wrap my head around that yet, so this conclusion is pending**). These initial results defied our predictions, suggesting that the availability of alternative demands erases preferences among demands (rather than inverting them).

@@

```{r Survival and prepost plots (Fig. 8), fig.width = 8, fig.height = 7, fig.align = "center", fig.cap = "Top: Survival curve (left), and distribution of quitting times over time (right). Bottom: Proportion accepted in the second half as a function of first half acceptances (left). Dashed lines indicate the mean for each cost. On the right, the proportion accepted every time participants experienced a given block type."}
# plots generated in 16.3.1. and 16.3.3.
(suppressWarnings(survPlot_wth) | suppressWarnings(RTplot_wth)) / (prepost_wth$scatterplot | prepost_wth$propBlocks)
```

  However, against our predictions, participants accepted significantly fewer trials on the second half of the experiment, regardless of cost (paired permutations; Cognitive: mean~pre~ = `r filter(df_mean, Cost == "Cognitive")$propAccept_1`, SD~pre~ = `r filter(df_mean, Cost == "Cognitive")$sdAccept_1`, mean~post~ = `r filter(df_mean, Cost == "Cognitive")$propAccept_2`, SD~post~ = `r filter(df_mean, Cost == "Cognitive")$sdAccept_2`, *p* `r report_p(prepost_wth$Perms[["Cognitive"]])`, Cohen's D = `r round(prepost_wth$CohenD[["Cognitive"]], digits = 2)`; Physical: mean~pre~ = `r filter(df_mean, Cost == "Physical")$propAccept_1`, SD~pre~ = `r filter(df_mean, Cost == "Physical")$sdAccept_1`, mean~post~ = `r filter(df_mean, Cost == "Physical")$propAccept_2`, SD~post~ = `r filter(df_mean, Cost == "Physical")$sdAccept_2`, *p* `r report_p(prepost_wth$Perms[["Physical"]])`, Cohen's D = `r round(prepost_wth$CohenD[["Physical"]], digits = 2)`; Wait-C:  mean~pre~ = `r filter(df_mean, Cost == "Wait-C")$propAccept_1`, SD~pre~ = `r filter(df_mean, Cost == "Wait-C")$sdAccept_1`, mean~post~ = `r filter(df_mean, Cost == "Wait-C")$propAccept_2`, SD~post~ = `r filter(df_mean, Cost == "Wait-C")$sdAccept_2`, *p* `r report_p(prepost_wth$Perms[["Wait-C"]])`, Cohen's D = `r round(prepost_wth$CohenD[["Wait-C"]], digits = 2)`; Wait-P: mean~pre~ = `r filter(df_mean, Cost == "Wait-P")$propAccept_1`, SD~pre~ = `r filter(df_mean, Cost == "Wait-P")$sdAccept_1`, mean~post~ = `r filter(df_mean, Cost == "Wait-P")$propAccept_2`, SD~post~ = `r filter(df_mean, Cost == "Wait-P")$sdAccept_2`, *p* `r report_p(prepost_wth$Perms[["Wait-P"]])`, Cohen's D = `r round(prepost_wth$CohenD[["Wait-P"]], digits = 2)`; Figure 10). The reduction in acceptances was not driven by a performance decline, as the difference in acceptance rates was not significantly correlated with the proportion of forced travels across individuals (Pearson's r = `r round(corDiff, digits = 2)`, *p* > 0.05). Notably, the pattern of choices from the first half of the experiment resembled that from the matching handling/travel block from the between-subject experiment (Figure 4A). Mixed effects logistic regressions performed on each half separately confirmed that the relative preference for cognitive effort seen previously was preserved at the beginning of the experiment (Figure 9, right, lower triangle), but that it disappeared over time amid an overall decline in acceptance rates as participants gained experience with all three conditions (upper triangle).

@@

```{r Prepost mixed effects plots (Fig. 9), fig.align = "center", fig.cap = "Acceptance rates over time. Left: Proportion accepted in the first two and last two blocks of the experiment. An initial preference for cognitive effort trials fades amid a global decrease in acceptance rates. Right: Mixed effects coefficient matrix for each half separately. The pattern from the first half resembles what was observed in the between-subjects experiment (Figure 4)."}
include_graphics("./Images/FIG10.png")

@@

``` {r BTW: 16.3.3. OC modeling, echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 4}
## fit the OC model
summaryOC <- list()
summaryOC$all <- dataBtw %>%
                 group_by(Cost, SubjID) %>%
                 do(optimizeOCModel(., simplify = T)) %>%
                 ungroup()

## plot
summaryOC$plot <- ggplot(summaryOC$all, aes(Cost, Gamma, fill = Cost)) +
  # geom_hline(yintercept = 0.64, alpha = 0.9, color = "gray20", size = 1) +
  # geom_hline(yintercept = 0.7, alpha = 0.9, color = "gray40", size = 1) +
  # geom_hline(yintercept = 0.9, alpha = 0.9, color = "gray60", size = 1) +
  geom_hline(yintercept = 0.74, alpha = 0.9, size = 1, linetype = "dashed") + # mean of highest earning rates across blocks
  #geom_dotplot(stackdir = 'center', binaxis = 'y', show.legend = F) +
  geom_boxplot(show.legend = F) +
  geom_jitter(width = 0.1, alpha = 0.5, show.legend = F, pch = 21, size = 3) +
  ylim(0,1.5) +
  labs(x = "") +
  scale_fill_manual(values = colsBtw) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        text = element_text(size = 16))

## one sample t against optimal strategy
summaryOC$t.test <- summaryOC$all %>%
  select(Cost, Gamma) %>%
  plyr::dlply("Cost", identity) %>%
  sapply(., function(data) {t.test(data$Gamma, mu = 0.74)})

## descriptive stats
summaryOC$descriptives <- summaryOC$all %>%
  group_by(Cost) %>%
  summarise(mGamma = round(mean(Gamma), digits = 2),
            sdGamma = round(sd(Gamma), digits = 2))

@@

``` {r BTW: 16.3.5. OC comparisons, echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 4}
# ANOVA
OCanova <- list()
OCanova$aov <- summary(aov(Gamma ~ Cost, data = summaryOC$all))
Fval <- round(OCanova$aov[[1]]$`F value`[1], digits = 2) # doing separate variables because I can't easily use the `` in markdown.
Pval <- OCanova$aov[[1]]$`Pr(>F)`[1]
Rsquared_aov <- round(OCanova$aov[[1]]$`Sum Sq`[1] / sum(OCanova$aov[[1]]$`Sum Sq`), digits = 2)

>>>>

---
title: Apparent preferences for cognitive effort fade when multiple forms of effort
  and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey^1^2, Gary Kane^1^2, & Joseph T. McGuire^1^2"
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
- \usepackage{indentfirst}
output:
  pdf_document:
    highlight: haddock
    keep_tex: no
  word_document: #default
    reference_docx: Style_reference_draft.docx
  html_document:
    df_print: paged
csl: apa.csl
indent: true
bibliography: cost_paper.bib
---

@@

```{r Trial plot (Fig. 1), eval = FALSE, out.width = "80%", fig.align = "center", fig.cap= "General foraging trial structure. On each trial, participants were offered to earn money (4, 8, or 20 cents) by sustaining effort or waiting during the handling time (2, 10, or 14 s). The end of a trial was followed by a travel time (handling and travel times always added up to 16 s). Participants could skip unfavorable trials and immediately start traveling to a potentially better offer. In the between-subject experiment, cost was fixed per participant and handling time varied per block. In the within-subject version, handling time was fixed at 10 s, but a combination of effort and delay trials changed per block. Possible costs, handling times, and rewards were fully disclosed in order to avoid experience-dependent learning."}

@@

  where *R* and *p* are the reward, and acceptance probability of offer i, respectively, *H* is the handling time for the block, and *T* its associated travel time. This gave us the total reward per second attainable in each block as a function of the acceptance threshold (i.e. the lowest amount accepted). Figure 2 shows possible earnings per second for each choice strategy, as well as the lowest amount participants should accept in order to maximize their rewards (circled dots). For example, a participant in a 10 s handling block should accept 8-cent and 20-cent rewards (and reject 4-cent rewards) to maximize their reward rate. Note that accepting every offer was often detrimental to participant earnings. Significant deviations from the optimal strategy can be understood as reflecting changes in subjective OC. This approach allowed us to test if certain demands boosted the value of offers, expressed as a subjective OC that fell below the optimal rate.

```{r Optimal plot (Fig. 2), eval = FALSE, out.width = "80%", fig.align = "center", fig.cap= "Possible earnings per second for each acceptance threshold (i.e. the smallest amount accepted) for each handling time. Circles denote the reward-maximizing threshold for each block, which is described in the table. Experiments shared the 10 s handling block, in which it was optimal to skip all 4 cent offers."}

@@

## MAYBE BETTER WAY TO DO PAIRWISE COMPARISONS, USING VECTORIZE FROM BELOW
# OCanova$perms <- summaryOC_btw$all %>%
#   select(Cost, Gamma) %>%
#   plyr::dlply("Cost", identity) %>%
#   do.call(cbind, .) %>%
#   select_if(is.numeric) %>%
#   rename(Physical = Physical.Gamma,
#          Cognitive = Cognitive.Gamma,
#          Wait = Wait.Gamma,
#          Easy = Easy.Gamma) %>%
#   outer(., ., FUN = Vectorize(permute), simple = T)
# diag(OCanova$perms) <- 1

@@

```{r General prop & earnings plots (Fig. 3), eval = FALSE, fig.width = 10, fig.height = 4, fig.align = "center", fig.cap = "Left: Proportion accepted per cost. Middle: Total number of dollars earned by each group by the end of the experiment. Right: The relationship between proportion accepted and total earned. Consistent with the foraging design, participants who over and under accepted earned the least."}
# plots generated in 16.2.1. and 16.1.3.
 prop1 | earnings1 | earnFitplot_btw
```

@@

```{r mixed effects plots (Fig. 4), eval = FALSE, out.width = "80%", fig.align = "center", fig.cap = "Comparing acceptance rates among costs for each reward and handling combination. A: Proportion accepted by each group for every combination of handling time and reward. Grey dots indicate the reward-maximizing acceptance for each combination. B: Matrix portraying the coefficients (color scale) and significance that resulted from switching the reference cost condition (crosses mark non-significant comparisons). Each entry shows how much more likely a group was to accept compared to the reference condition (row). C: AIC and BIC values for a forward mixed-effects model comparison, ranging from a base intercept-only to a full-interaction model."}
#hrPlot_btw / (ABIC_plot_btw | ME_matrix_btw) # doesn't work with corrplot.. oh well
include_graphics("./Images/FIG4.png")
```

@@

```{r Survival and prepost plots (Fig. 5), eval = FALSE, fig.width = 8, fig.height = 7, fig.align = "center", fig.cap = "Top: Survival curve (left), and distribution of times when participants quit a trial over time. Bottom: Proportion accepted in the second half as a function of first half acceptances (left). Dashed lines indicate the mean for each group. On the right, the proportion accepted over time per group. The order of blocks was repeated in the second half (indicated by the black dashed line)."}
# plots generated in 16.3.1. and 16.3.2.
(suppressWarnings(survPlot_btw) | RTplot_btw) / (prepost_btw$scatterplot | prepost_btw$propBlocks)
```

@@

```{r Prepost mixed effects plot (Fig. 6), eval = FALSE, fig.align = "center", fig.cap = "Distilling acceptance behavior before and after the break. Left: Proportion of acceptances per handling, reward, and half, showing that effortful demands became increasingly aversive over time. Right: Mixed-effects coefficients denoting the comparison among costs acceptances for each half separately (mirrored along the diagonal). Even though acceptances steadily decreased in the effort groups, the relative preferences were preserved."}
include_graphics("./Images/FIG6.png")

@@

  Analyses from the previous experiment were repurposed to examine cost differences within individuals. We started by testing the hypothesis that participants would successfully integrate the statistics of the environment. A mixed-effects logistic regression once again showed that larger reward amounts significantly increased acceptance probabilities ($\beta$ = `r historyCoeffs_wth["Offer", "Estimate"]`, SE = `r historyCoeffs_wth["Offer", "Std. Error"]`, *p* `r report_p(historyCoeffs_wth["Offer", "Pr(>|z|)"])`), and that participants became more selective if they had recently observed large offers ($\beta$ = `r historyCoeffs_wth["recentOffers", "Estimate"]`, SE = `r historyCoeffs_wth["recentOffers", "Std. Error"]`, *p* `r report_p(historyCoeffs_wth["recentOffers", "Pr(>|z|)"])`). Model comparisons suggested that participants were influenced by the `r historyIndex_wth + 1` most recent rewards (AIC = `r round(AIC(historySize_wth[[historyIndex_wth]]), digits = 2)`). However, this time participants were unaffected by recent reward misses (*p* > 0.05).
  As before, a linear model predicting earnings showed that participants who completed too few or too many trials earned the least, in line with the experimental design (Figure 7A; general linear model with quadratic term, F = `r round(summary(lmEarn)$fstatistic[1], digits = 2)`, $\beta$ = `r round(summary(lmEarn)$coefficients[3,1], digits= 2)`, SE = `r round(summary(lmEarn)$coefficients[3, 2], digits = 2)`, $R^2$ = `r round(summary(lmEarn)$r.squared, digits = 2)`, *p* `r report_p(summary(lmEarn)$coefficients[3, 4])`). When assessing the optimality of acceptances for 4 and 8 cents (99% of 20 cent offers were correctly accepted), chi-squared tests against the optimal rate of 50% showed that participants under-accepted physical effort trials (proportion accepted = `r round(summary_costvsreward["Physical", "probability"], digits = 2)`, $\chi^2$ = `r round(summary_costvsreward["Physical", "chisquared"], digits = 2)`, *p* `r report_p(summary_costvsreward["Physical", "pval"])`; all other *p* > 0.05). Despite this deviation, participants earned a similar total amount per cost on average (mean Cognitive = `r filter(earnings_within, Cost == "Cognitive")$mEarn`, SD = `r filter(earnings_within, Cost == "Cognitive")$sdEarn`; mean Wait-C = `r filter(earnings_within, Cost == "Wait-C")$mEarn`, SD = `r filter(earnings_within, Cost == "Wait-C")$sdEarn`; mean Physical = `r filter(earnings_within, Cost == "Physical")$mEarn`, SD = `r filter(earnings_within, Cost == "Physical")$sdEarn`; mean Wait-P = `r filter(earnings_within, Cost == "Wait-P")$mEarn`, SD = `r filter(earnings_within, Cost == "Wait-P")$sdEarn`). These results convey that participants in both experiments were similarly influenced by the characteristics of the environment, although participants in the present experiment gravitated more towards optimality.

@@

```{r Earnings/prop completed/ME plots (Fig. 7), eval = FALSE, fig.width = 8, fig.height = 7, fig.align = "center", fig.cap = "A: Quadratic relationship between trial completions and total earnings. B: proportion of acceptances of each reward per cost condition. C: Matrix of mixed-effects coefficients showing comparisons among costs (while controlling for reward amount). Comparisons between effort and wait trials from different blocks are omitted, as these are not considered in our original predictions. Crosses mark non-significant comparisons. D: AIC and BIC values for models of increasing complexity (a priori = AllMain)."}

@@

  Next, we addressed the hypothesis that passive waiting trials would be less costly than effortful ones. Even though the reward acceptance pattern at 10 s handling was similar to the between-subjects experiment (Figure 7B), differences among cost conditions were less evident. Moreover, these differences contradicted our predictions, as cognitive effort remained the least costly among demands. A mixed effects logistic regression confirmed once again that increases in offer amount made acceptances significantly more likely (global intercept = `r round(fixef(mixLogis_main$Cognitive)["(Intercept)"], digits = 2)`, $\beta$ = `r round(fixef(mixLogis_main$Cognitive)["Offer"], digits = 2)`, SE = `r round(summary(mixLogis_main$Cognitive)[[10]]["Offer", "Std. Error"], digits = 2)`, *p* `r report_p(summary(mixLogis_main$Cognitive)[[10]]["Offer", "Pr(>|z|)"])`), but that the only significant difference was the higher likelihood to accept cognitive offers compared to physical ones (Figure 7C; $\beta$ = `r round(mixSummary_wth$Betas["Physical", "Cognitive"], digits = 2)`, *p* `r report_p(mixSummary_wth$Pvals["Physical", "Cognitive"])`). Model comparisons suggested that even though a model with all main effects explained a large portion of the variance ($R^2$ = `r round(mixLogis_compare$Rsq, digits = 2)`), there were similar AIC and BIC values for any model including reward amount (Figure 7D). An analysis of deviance among the three models showed that adding cost condition to reward amount significantly improved fit (AIC~reward~ = `r filter(abicLogis_compare, Model == "Reward")$AIC`, AIC~a-priori~ = `r filter(abicLogis_compare, Model == "AllMain")$AIC`, $\chi^2$(`r mixLogis_compare$aod[[7]][2]`) = `r round(mixLogis_compare$aod$Chisq[2], digits = 2)`, *p* `r report_p(mixLogis_compare$aod[[8]][2])`). However, probing the interaction between these features significantly improved fits from the main effects model (AIC~Interaction~ = `r filter(abicLogis_compare, Model == "Interaction")$AIC`, $\chi^2$(`r mixLogis_compare$aod[[7]][3]`) = `r round(mixLogis_compare$aod$Chisq[3], digits = 2)`, *p* `r report_p(mixLogis_compare$aod[[8]][3])`). (**discuss with Joe: the interaction flips results over, such that both waits > pysical effort = cognitive effort...as predicted originally? Can't wrap my head around that yet, so this conclusion is pending**). These initial results defied our predictions, suggesting that the availability of alternative demands erases preferences among demands (rather than inverting them).

@@

```{r Survival and prepost plots (Fig. 8), eval = FALSE, fig.width = 8, fig.height = 7, fig.align = "center", fig.cap = "Top: Survival curve (left), and distribution of quitting times over time (right). Bottom: Proportion accepted in the second half as a function of first half acceptances (left). Dashed lines indicate the mean for each cost. On the right, the proportion accepted every time participants experienced a given block type."}
# plots generated in 16.3.1. and 16.3.3.
(suppressWarnings(survPlot_wth) | suppressWarnings(RTplot_wth)) / (prepost_wth$scatterplot | prepost_wth$propBlocks)
```

  However, against our predictions, participants accepted significantly fewer trials on the second half of the experiment, regardless of cost (paired permutations; Cognitive: mean~pre~ = `r filter(df_mean, Cost == "Cognitive")$propAccept_1`, SD~pre~ = `r filter(df_mean, Cost == "Cognitive")$sdAccept_1`, mean~post~ = `r filter(df_mean, Cost == "Cognitive")$propAccept_2`, SD~post~ = `r filter(df_mean, Cost == "Cognitive")$sdAccept_2`, *p* `r report_p(prepost_wth$Perms[["Cognitive"]])`, Cohen's D = `r round(prepost_wth$CohenD[["Cognitive"]], digits = 2)`; Physical: mean~pre~ = `r filter(df_mean, Cost == "Physical")$propAccept_1`, SD~pre~ = `r filter(df_mean, Cost == "Physical")$sdAccept_1`, mean~post~ = `r filter(df_mean, Cost == "Physical")$propAccept_2`, SD~post~ = `r filter(df_mean, Cost == "Physical")$sdAccept_2`, *p* `r report_p(prepost_wth$Perms[["Physical"]])`, Cohen's D = `r round(prepost_wth$CohenD[["Physical"]], digits = 2)`; Wait-C:  mean~pre~ = `r filter(df_mean, Cost == "Wait-C")$propAccept_1`, SD~pre~ = `r filter(df_mean, Cost == "Wait-C")$sdAccept_1`, mean~post~ = `r filter(df_mean, Cost == "Wait-C")$propAccept_2`, SD~post~ = `r filter(df_mean, Cost == "Wait-C")$sdAccept_2`, *p* `r report_p(prepost_wth$Perms[["Wait-C"]])`, Cohen's D = `r round(prepost_wth$CohenD[["Wait-C"]], digits = 2)`; Wait-P: mean~pre~ = `r filter(df_mean, Cost == "Wait-P")$propAccept_1`, SD~pre~ = `r filter(df_mean, Cost == "Wait-P")$sdAccept_1`, mean~post~ = `r filter(df_mean, Cost == "Wait-P")$propAccept_2`, SD~post~ = `r filter(df_mean, Cost == "Wait-P")$sdAccept_2`, *p* `r report_p(prepost_wth$Perms[["Wait-P"]])`, Cohen's D = `r round(prepost_wth$CohenD[["Wait-P"]], digits = 2)`; Figure 9). The reduction in acceptances was not driven by a performance decline, as the difference in acceptance rates was not significantly correlated with the proportion of forced travels across individuals (Pearson's r = `r round(corDiff, digits = 2)`, *p* > 0.05). Notably, the pattern of choices from the first half of the experiment resembled that from the matching handling/travel block from the between-subject experiment (Figure 4A). Mixed effects logistic regressions performed on each half separately confirmed that the relative preference for cognitive effort seen previously was preserved at the beginning of the experiment (Figure 9, right, lower triangle), but that it disappeared over time amid an overall decline in acceptance rates as participants gained experience with all three conditions (upper triangle).

@@

```{r Prepost mixed effects plots (Fig. 9), eval = FALSE, fig.align = "center", fig.cap = "Acceptance rates over time. Left: Proportion accepted in the first two and last two blocks of the experiment. An initial preference for cognitive effort trials fades amid a global decrease in acceptance rates. Right: Mixed effects coefficient matrix for each half separately. The pattern from the first half resembles what was observed in the between-subjects experiment (Figure 4)."}
include_graphics("./Images/FIG10.png")

@@

``` {r BTW: 16.3.3. OC modeling, echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 4}
## fit the OC model
summaryOC_btw <- list()
summaryOC_btw$all <- dataBtw %>%
                 group_by(Cost, SubjID) %>%
                 do(optimizeOCModel(., simplify = T)) %>%
                 ungroup()

## plot
summaryOC_btw$plot <- ggplot(summaryOC_btw$all, aes(Cost, Gamma, fill = Cost)) +
  # geom_hline(yintercept = 0.64, alpha = 0.9, color = "gray20", size = 1) +
  # geom_hline(yintercept = 0.7, alpha = 0.9, color = "gray40", size = 1) +
  # geom_hline(yintercept = 0.9, alpha = 0.9, color = "gray60", size = 1) +
  geom_hline(yintercept = 0.74, alpha = 0.9, size = 1, linetype = "dashed") + # mean of highest earning rates across blocks
  #geom_dotplot(stackdir = 'center', binaxis = 'y', show.legend = F) +
  geom_boxplot(show.legend = F) +
  geom_jitter(width = 0.1, alpha = 0.5, show.legend = F, pch = 21, size = 3) +
  ylim(0,1.5) +
  labs(x = "") +
  scale_fill_manual(values = colsBtw) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        text = element_text(size = 16))

## one sample t against optimal strategy
summaryOC_btw$t.test <- summaryOC_btw$all %>%
  select(Cost, Gamma) %>%
  plyr::dlply("Cost", identity) %>%
  sapply(., function(data) {t.test(data$Gamma, mu = 0.74)})

## descriptive stats
summaryOC_btw$descriptives <- summaryOC_btw$all %>%
  group_by(Cost) %>%
  summarise(mGamma = round(mean(Gamma), digits = 2),
            sdGamma = round(sd(Gamma), digits = 2))

@@

``` {r BTW: 16.3.5. OC comparisons, echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 4}
# ANOVA
OCanova <- list()
OCanova$aov <- summary(aov(Gamma ~ Cost, data = summaryOC_btw$all))
Fval <- round(OCanova$aov[[1]]$`F value`[1], digits = 2) # doing separate variables because I can't easily use the `` in markdown.
Pval <- OCanova$aov[[1]]$`Pr(>F)`[1]
Rsquared_aov <- round(OCanova$aov[[1]]$`Sum Sq`[1] / sum(OCanova$aov[[1]]$`Sum Sq`), digits = 2)
